Virtual Register Scheduling

### 2014-12-17 VEE-15/paper.tex
old thing

% Make text in figure captions small
\usepackage{caption3} % load caption package kernel first
\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
\usepackage[small]{caption}

@@

% Compressed enumerations
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{2pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}

@@

% For writing Granary tool names.
\newcommand{\toolname}[1]{{\scshape #1}}

@@

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{Fast and Flexible Dynamic Binary Translation for the Kernel}

@@

\maketitle
\subsection*{Abstract}
TODO

\section{Introduction}\label{sec:intro}

Granary is a framework for creating flexible and efficient tools that instrument the Linux kernel. Granary's key novelty is its \emph{runtime code specialization} feature, which enables tools to decide what code to instrument, and how to instrument that code. Flexible specialization allows tools to be picky about instrumentation. For example, a tool can choose to instrument every load and store operation, but only inside nested critical sections that are executing in module code. Granary's runtime code specialization feature addresses two limitations of existing kernel analysis systems: \begin{inparaenum}[i)]
	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU}; and
	\item they can only instrument code using a single policy.
\end{inparaenum}

Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analyzed (e.g., modules, functions). In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code. Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the baseline overhead of the runtime system supporting that instrumentation. Also, specializing based on what is instrumented allows instrumentation tools to focus instrumentation on the code of interest. For example, if one wants to instrument RCU read-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out and find those critical sections is wasteful.

Another limitation of existing systems is that they are unable to perform efficient runtime code specialization. That is, they are unable to change how code is instrumented based on the context in which that code will run. This is unfortunate because it means that tools wanting to instrument code differently must create their own solution that works around this deficiency. For example, one way that existing tools can specialize code is to generate as many versions of the code as there are potential specializations. When instrumented code is executing, runtime checks continually check and jump to the correct version of the code for the current execution context. However, runtime checks are inefficient, and generating many versions of the code introduces bloat and pollutes the CPU's instruction cache. Granary avoids these issues by supporting context-specific runtime code specialization that generates multiple versions of basic blocks \emph{just-in-time} (JIT), instead ahead-of-time, and without requiring runtime checks.

One example tool that uses Granary's JIT-based runtime code specialization feature is \toolname{RCUdbg}: Granary's Read-Copy-Update (RCU) debugging tool. \toolname{RCUdbg} catches bugs related to invalid uses of RCU-protected pointers (\Cref{sec:rcudbg}). \toolname{RCUdbg} distinguishes between two kinds of RCU-protected pointers (reader-side and writer-side) and plain old pointers, and three execution contexts (inside of a read-side critical section, inside of a write-side critical section, and in non-RCU-related code). The actions taken when a pointer is de-referenced varies depending on the kind of pointer and the execution context. Code executing in read-side critical sections is specialized to enforce that only plain-old and read-side pointers are accessed, and that all accesses through read-side pointers are memory reads. Write-side critical sections are specialized in a similar way.

The \toolname{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code executing outside of both contexts runs natively or with lightweight instrumentation. Achieving this level of specialization using and existing system requires complicated runtime state management and checks to switch between different versions of the code. With Granary, a tool like \toolname{RCUdbg} can specialize code using \emph{instrumentation policies}. An instrumentation policy is a function that names an execution context and decides how to instrument instructions that will execute within that context. When a policy is used to instrument a sequence of instructions (called a basic block), a new version of the basic block is generated and cached.

\toolname{RCUdbg} defines three policies, and declares transitions between these policies when control flows in and out of RCU read- and write-side critical sections. When a function like \texttt{memset} is invoked in one context, the policy associated with that context is used to generate an instrumented version of \texttt{memset} for that context. If, for example, \texttt{memset} is never invoked within a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context. Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated.

The key insight of Granary's approach to runtime code specialization is that code versioning is a context-tracking mechanism. That is, if an instruction from a basic block instrumented by \toolname{RCUdbg}'s read-side critical section policy is being executed, then the current thread must be executing inside of a read-side critical section. Similarly, if the read-side policy is used to instrument a basic block, then \texttt{RCUdbg} knows at instrumentation time\footnote{Instrumentation time is analogous to ``compile time", in that a policy is generally used to instrument a basic block of code once, and then that basic block may be executed many times.} that this basic block will only execute within a read-side critical section. If the basic block contains a \texttt{call} to the \texttt{memset} function, then \texttt{memset} is also known to execute within a read-side critical section because it does not terminate a read-side critical section. By deciding which version of \texttt{memset} to execute at instrumentation time, \toolname{RCUdbg} avoids expensive runtime checks that other systems impose in order to achieve similar kinds of specialization. \toolname{RCUdbg} can specify that \texttt{memset} will also execute within a read-side critical section by telling Granary that it should use the read-side critical section policy to instrument \texttt{memset}.

The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how a Granary tool can choose to instrument only specific code, \Cref{sec:how} describes how Granary change the policy used to instrument code, etc.

\section{Dynamic Binary Translation}\label{sec:dbt}

Granary uses dynamic binary translation (DBT) to instrument the Linux kernel. DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimization}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}). We chose DBT because \begin{inparaenum}[i)]
	\item static binary analysis tools are unable to cope with \texttt{x86-64}'s mix of variably-sized instructions and data; 
	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
	\item virtualization-based approaches are unable to analyze non-virtualizable device drivers \cite{DRK}.
\end{inparaenum}

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch (\texttt{jCC}), \texttt{ret}, or a \texttt{jmp} instruction. Unlike DRK and btkernel, Granary does not terminate basic blocks at \texttt{call} instructions. This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}). Granary takes control of execution starting from ``natural" program entrypoints (e.g. interrupt vectors, system call entrypoints, module initializers). Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as control flow reaches them for the first time.

Translated basic blocks are linked together and stored in a \emph{code cache}. Basic blocks are indexed by $(pc, policy)$ pairs, where $pc$ is the native program counter associated with the beginning of the basic block's instructions, and $policy$ is the instrumentation policy used to instrument the native instructions starting at $pc$. The index is maintained by a globally accessible hash table.

Granary's JIT-based translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute. This happens when there is a control flow transfer in a basic block, and the target of that control flow transfer is not already indexed in Granary's code cache. When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates. Granary operates on a CPU-private stack for two reasons. First, if Granary did not switch stacks, then context switches into Granary risk overflowing the runtime call stack. This can happen because kernel stacks are typically small (4K to 8K in size), and because a context switch could happen when the native stack is nearly exhausted. Second, on an SMP, multiple processors can simultaneously execute Granary code (e.g. to translate code that will later execute). In such cases, Granary depends on CPU-private data structures to avoid contention, and the private stack is one such data structure. A context-switch back to the code cache occurs when the next basic block has been found or translated. At this point, instrumented execution continues. Similar to other DBT systems, Granary uses hot code patching (also called code splicing) to reduce the number of context switches. In our experience, kernel code stabilizes very quickly: context switches stop happening after the first few seconds of translation.

\subsection{Basic Block Meta-Data}\label{sec:metadata}

Associated with each basic block is meta-data that describes the basic block. Basic block meta-data includes: \begin{itemize}
	\item The length in bytes of the original and translated basic blocks, and the address of the first instruction in the original basic block and the instrumented basic block. This information helps provide a rich debugging interface for instrumentation tool authors. Granary is unique among DBT tools in that the debugging experience is actually pleasant. For example, Granary has a built-in reverse debugging mechanism, which allows developers to inspect previously executed basic blocks and past register state.
	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block. This kernel-specific meta-data is used by Granary to implement transparent page fault exception handling (\Cref{sec:interrupt_return_address_transparency}).
	\item Ranges of instructions that must be executed atomically. Granary uses this information to implement selective interrupt delaying (\Cref{sec:interrupt_delay}).
	\item The allocator used to allocate the instructions of the basic block. Granary uses this information to implement its trace-allocation optimization (\Cref{sec:trace_alloc}).
	\item The instrumentation policy used to instrument the basic block. Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines. For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
\end{itemize}

Basic block meta-data can also be extended by instrumentation tools. This makes developing instrumentation tools that record per-basic block statistics very easy to make. For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts (\Cref{sec:cfg}). This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization. This meta-data is specified by \toolname{cfg} and automatically allocated by Granary.

\subsection{Transparency}\label{sec:transparency}
An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, the only artifact Granary exposes is that interrupt and function return addresses can point into the code cache.

\subsubsection{Function return addresses}\label{sec:return_address_transparency}
Unlike most DBT frameworks, Granary's basic blocks do not terminate at \texttt{call} instructions. This exposes code cache addresses to instrumented module code in the form of return addresses stored on the runtime call stack. The benefit of including \texttt{call} instructions are: \begin{enumerate}

	\item {\bf Zero overhead procedure returns.} Unlike Granary, DRK emulates procedure calls and returns. DRK emulates \texttt{call} instructions by pushing the intended native return address onto the runtime call stack, followed by a direct direct jump to the next (instrumented) basic block \cite{DRK}. This approach enables return address transparency, but introduces additional overheads on procedure returns, which must do costly indirect branch lookups to find the intended return target.

Btkernel's approach is similar in spirit to Granary's; however, btkernel ends basic blocks at \texttt{call} instructions. This introduces unnecessary control-flow on the return path from procedures. That is, btkernel's ``call/ret optimization" ostensibly enables zero-overhead procedure returns by leaving \texttt{ret} instructions as-is; however, the benefits are more apparent than real. To transfer control to the next logical basic block after a procedure return, btkernel's \texttt{ret}s first transfer control to the end of the basic block that performed the \texttt{call}, and then \texttt{jmp} to the next basic block. Granary avoids this unnecessary \texttt{jmp} because basic blocks do not end at \texttt{call} instructions.

In some cases, emulating DRK's return-address lookup mechanism is desirable. For example, if Granary attaches while some program's execution is underway (and so the runtime call stack contains native return addresses), then a direct \texttt{ret} back to one of the native return addresses might be undesirable (e.g. if the tool wants comprehensive instrumentation of all code). In such cases, a tool can partially specialize instrumentation according to whether or not the return address might be in the code cache. Partial specialization is enabled by \emph{policy properties}, which are transparently inherited along with normal instrumentation policy information. In this case, the policy property can be used to tell Granary to emulate \texttt{ret} instructions when the return address is potentially a native address, and directly \texttt{ret} when the return address is known to be in the code cache.\footnote{This feature is used when Granary is configured to instrument user space programs. User space Granary depends on the Linux dynamic loader to inject Granary into a running process. When injected, a Granary takeover function is executed on an already active native runtime call stack. Granary's takeover procedure transparently emulates returns, thus taking over the previously fully-native call stack. This feature allows instrumentation tools to focus on their particular objectives and not the intricacies of thread takeover.}

	\item {\bf Improved visibility.} Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread. \toolname{RCUdbg} benefits from this additional visibility because it can observe the end of read-side critical sections (\texttt{call}s to \texttt{rcu\_read\_unlock}), and optimistically assume that functions executing before the read-side critical section ends will execute within the read-side critical section.

\end{enumerate}

\subsubsection{Interrupt return addresses}\label{sec:interrupt_return_address_transparency}
Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts. A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}. The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the exception return address belongs to one of the pre-defined ranges. Granary records any exception table entries as part of each basic block's meta-data, so that a page fault in a basic block can be mapped to the correct native faulting program counter.

\subsection{Reentrancy}
Reentrancy is an important concern for all DBT systems, and is especially important for DBT systems operating in the kernel environment. For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state \cite{DRK}. Preserving interrupt transparency, however, requires that interrupts either be delayed until the end of any injected instrumentation instructions. Together, these requirements make it very difficult to write instrumentation tools.

\subsubsection{Granary}
For efficiency, most kernel-space instrumentation systems (Granary included) operate on CPU-private state (e.g. Granary's CPU-private stacks). However, accessing CPU-private data without first disabling interrupts results in undefined behavior. Therefore, context-switches into Granary first disable interrupts and then switch execution onto one of Granary's CPU-private stacks. Context switches out of Granary and into native code or the code cache returns execution to the native stack and re-enable interrupts. Execution behavior is preserved because Granary saves the machine state (i.e. registers, flags) in a reentrant way. All other parts of Granary are fully reentrant.

\subsubsection{Instrumentation}\label{sec:interrupt_delay}

\begin{figure}[t!]
\begin{multicols}{2}
\vspace{-1.5em}\subfloat[\label{fig:interrupt_delay_drk}Pre and post instrumentation in DRK. Injected instrumentation blocks are treated as atomic with respect to interrupts.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/drk_instrumented_mov.pdf}}\vspace{-2.5em}
\columnbreak
\vspace{-1.5em}\hspace{-0.5em}\subfloat[\label{fig:interrupt_delay_granary}Pre and post instrumentation in Granary. The two instrumentation blocks, as well as the native instruction are wrapped in a delay region.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/granary_instrumented_mov.pdf}}\vspace{-2.5em}
\end{multicols}
\caption{Example showing equivalent pre- and post-instruction instrumentation with DRK (left) and Granary (right). In DRK, the injected instrumentation instructions (grey) are guaranteed to execute atomically, up until the next native instruction (white). In Granary, no guarantees are made on injected instrumentation, unless otherwise instructed. In this case, a delay region surrounds both the pre- and post-instrumentation, as well as the native instruction, guaranteeing that the entire sequence of instructions execute on the same CPU.}
\end{figure}

One challenge with any DBT system is how it copes with non-reentrant instrumentation. For example, instrumentation that modifies CPU-private data structures (without disabling interrupts) cannot be safely executed by a pre-emptive kernel running on an SMP. If the thread being instrumented is interrupted, then the kernel might decide to re-schedule the thread to resume its execution on a different CPU. Any modifications to CPU-private state would lead to undefined behavior. An example of this is instrumentation that acquires a CPU-private lock on one CPU, is interrupted, and later releases the wrong lock on the wrong CPU.

DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}. Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86-64: interrupts and exceptions only arrive on native instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries. This approach introduces subtle complexity when injecting non-idempotent instrumentation before (pre) and after (post) native instructions.

\Cref{fig:interrupt_delay_drk} shows an example of pre- and post-instrumentation around a \texttt{mov} instruction. If an interrupt arrives within the pre-instrumentation, then DRK delays they delivery of the interrupt until the native instruction (\texttt{mov}). After the interrupt is handled and execution attempts to return to the interrupted instructions, DRK will re-instrument the ``tail" of interrupted basic block, starting from the \texttt{mov} instruction. This involves re-creating the pre- and post-instrumentation of the \texttt{mov} in the tail basic block. When executed, this will result in an unbalanced number of \texttt{push} and \texttt{pop} instructions. That is, \texttt{push \%rax} is executed in both the original and tail basic block, but \texttt{pop \%rax} is only executed in the tail basic block. This is a specific example of a more general problem: DRK's approach to interrupt delaying does not play well with non-idempotent instrumentation.

Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}. The is inefficient because it introduces flag stalls into the instruction pipeline \cite{AgnerMicroarchitecture}, and because interrupts arrive relatively infrequently.

Granary charts a middle ground between the two aforementioned approaches. Most instrumentation is re-entrant, partially because kernel instrumentation tools can depend on the runtime call stack for scratch storage space. However, interrupt delaying is sometimes useful because it enables non-reentrant instrumentation to avoid having to continually disable and re-enable interrupts. For example, Granary's \toolname{Guardrail} tool records traces of executed module instructions into CPU-private buffers \cite{Guardrail}. These traces are concurrently analyzed for bugs by another thread. Operating on a CPU-private buffer is more efficient than operating on a global buffer because it avoids cache line contention and locking on an SMP. However, writing to a CPU-private buffer is only safe if the entire write executes on the same CPU. An equivalent tool developed for btkernel---which does not support interrupt delaying---would have to disable and the re-enable interrupts around the write.

Granary conveniently solves this problem by allowing instrumentation tools to declare \emph{interrupt delay regions}. An interrupt delay region is a sequence of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically. \Cref{fig:interrupt_delay_granary} shows an interrut delay region that surrounds both the pre- and post-instrumentation of a native \texttt{mov} instruction. If an interrupt arrives anywhere within this region of code, Granary will delay the interrupt to the end of the region. The declarative nature of interrupt delay regions allows instrumentation developers to opt-in to interrupt delaying, and control exactly when, where, and for how long interrupts can be delayed.

Granary implements interrupt delay regions by recording bounds information about interrupt delay regions inside of each basic block's meta-data (\Cref{sec:metadata}). When a basic block is interrupted, Granary inspects the block's meta-data and determines if the interrupt must be delayed. If a delay is required, then all instructions belonging to the delay range are copied into a CPU-private buffer. All instructions are copied because a delay region might contain internal control-flow. A CPU-private buffer is used because the number of CPUs is a natural upper bound on the number of simultaneously executing delay regions. After the copied instructions, Granary emits code that code re-creates the interrupt stack frame, followed by a jump to the interrupt vector (which is controlled by Granary). This jump ensures that Granary's interrupt handler regains control after the instructions within the delay region fully execute, but before the kernel observes the interrupt. When Granary regains control, it can then delegate to the instrumentation tool to handle the interrupt, or defer handling of the interrupt to the kernel.

When all instructions have been emitted to the CPU-private buffer, Granary emulates a return from the interrupt (without issuing an \texttt{iret} instruction). The emulated return leaves interrupts disabled and transfers control to the CPU-private buffer. Leaving interrupts disabled ensures that the remainder of the delay region will execute on the current CPU, and avoids reentrancy issues related to the copied region being located in a CPU-private buffer.

\subsection{Control Flow Implementation}

Granary's mechanism of looking up branch targets and resolving control flow depends only on control flow instruction (CTI) policy and target information. This information is either directly embedded in emitted executable code, or in immutable data structures directly referenced by emitted code. This approach ensures that policy information is never lost or corrupted, even in the face of concurrent executions of the same emitted code, arbitrary pre-emption, and arbitrary resumption. Because Granary's translation mechanism is driven by these two inputs, Granary is otherwise stateless\footnote{This is not quite true, as Granary maintains persistent state in the form of the code cache index. However, the code cache index behaves similarly to memoization in a purely functional programming language: it is an optimization, not a necessity.} across context switches.

\subsubsection{Direct Control Flow}\label{sec:direct_control_flow}
\lstset{language=[x64]Assembler}
\newsavebox\nativejcc
\begin{lrbox}{\nativejcc}
\begin{minipage}[b]{4cm}
\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
cmp $0x0, %rax
jz value_is_zero 
...
( fall through )
\end{lstlisting}
\end{minipage}
\end{lrbox}

\newsavebox\translatedjcc
\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
  cmp $0x0, %rax
  jz edge_value_is_zero
  jmp edge_fall_through

edge_value_is_zero:
  call *(jz_patch_data)
edge_fall_through:
  call *(jmp_patch_data)
\end{lstlisting}
\end{minipage}
\end{lrbox}

\begin{figure}[t!]
\subfloat[Native Code]{\usebox\nativejcc}
\hfill
\subfloat[Translated Code]{\usebox\translatedjcc}
\caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code. The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_value\_is\_zero}). When executed, the function call reads its target from an edge data structure (\texttt{jz\_patch\_data}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}. Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction. Finally, the return address of the \texttt{call} instruction is altered to transfer control to the newly resolved target.}
\end{figure}
Like btkernel, Granary translates direct CTIs into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}). Edge code lives outside the code cache, so as to not pollute the processor's instruction cache \cite{btkernel}.

When control flows through a basic block $B_{pred}$ for the first time, any jumps taken will transfer control to edge code that is specific to the jump taken. Edge code contains only a single indirect \texttt{call} instruction, thus allowing multiple edge codes to occupy a single cache line. The indirect \texttt{call} transfers control to Granary via a function pointer located in an ``edge data structure", which describes how to translate the target basic block, $B_{succ}$. Granary gains control and looks up or translates the target code according instrumentation policy stored in the edge data structure. Then, Granary modifies the original CTI in $B_{pred}$ so that it transfers control to $B_{succ}$  (\Cref{sec:patching}). Finally, Granary releases control and context switches to $B_{succ}$.

\subsubsection{Indirect Control Flow}\label{sec:ibl}

Granary translates indirect CTIs (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions branch to ``in-edge" code. In-edge code loads the indirect jump target (e.g. a specific register or memory location), saves the flags register, and several scratch registers. Finally, the in-edge uses the indirect target as an index into a global, fixed-size hash table of indirect branch targets. The entries of the hash table either point to Granary or to ``out-edge" code (\Cref{fig:ibl}).

\begin{figure}[t!]
\hspace{4em}\includegraphics[height=13em,clip=true,trim=0 5em 10em 0]{diagrams/ibl.pdf}
\caption{\label{fig:ibl}Process of resolving an indirect branch. Control transfers from a source basic block, to the in-edge code specific to where the indirect target is stored (e.g. register, memory address), through the global hash table and on to either one or more out-edge code stubs, or to Granary.}
\end{figure}

Out-edge code is specific to a target basic block, and performs the function of separate chaining in the indirect branch lookup hash table. When control transfers to out-edge code, the out-edge checks to see if the indirect target matches the out-edge code's target basic block. If they match, then the out-edge code restores scratch and flags registers and jumps directly to the target basic block. Otherwise, the out-edge code jumps to another out-edge code or to Granary. If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table. In this respect, out-edge code is chained together through control flow, just as colliding entries in a typical hash table can be chained together through a linked list.

Granary's use of out-edge code avoids re-entrancy issues related to indirect control flow that are present in DRK and btkernel. Both DRK and btkernel store the target of an indirect CTI in a CPU-private memory location. To complete an indirect control transfer, both systems perform (roughly) the following steps: \begin{enumerate}[itemindent=\parindent]%
\setlength{\itemsep}{2pt}%
\setlength{\parskip}{0pt}%
	\item Save the flags and disable interrupts.
	\item Save registers.
	\item Lookup the next basic block address.
	\item Store that address in a CPU private memory location.
	\item Restore registers.
	\item Restore the flags and interrupt state.
	\item Indirectly jump to the target basic block through the CPU-private location.
\end{enumerate} %
However, the ordering of steps (6) and (7) introduce a reentrancy problem: the access to the CPU-private location is not safe because an interrupt might arrive before the indirect jump in step (7). This unusual design is motivated by the constraint that the entire lookup procedure must save (2) and restore (5) the register state for correctness, but must also save the target pc \emph{somewhere} so that execution can eventually be redirected. Granary's out-edge code avoids this reentrancy issue by having the out edge code itself (in the form of a direct \texttt{jmp}) contain the target pc.

\subsection{Optimizations}\label{sec:optimize}

\subsubsection{Ahead-of-time Translation Optimization} \label{sec:aot}
Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s. Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e. when the conditional jump is not taken).

If a basic block ends in either a copy of a native direct \texttt{jmp} or synthesized direct \texttt{jmp}, then Granary will perform ahead-of-time translation and instrumentation of the targeted basic block. The repeated application of translating \texttt{jmp} targets forms traces in Granary's code cache. Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding \texttt{jmp}s between adjacent basic blocks.

Granary's tracing strategy is based on the observation that the most commonly executed path through code follows a straight line. That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI. We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's internal performance counters to count how many conditional branches were patched (by their edge code, \texttildelow 35\%), versus how many synthesized fall-through \texttt{jmp}s were patched (\texttildelow 80\%).

Profile-guided ahead-of-time tracing further improves on Granary's otherwise na\"{i}ve tracing strategy by integrating profile information from Granary's \toolname{cfg} tool (\Cref{sec:cfg}).

\subsubsection{Patching and Alignment  Optimizations}\label{sec:patching}
For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}). That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch in the translated code.

Granary opportunistically connects two basic blocks together using a direct branch when: \begin{inparaenum}[i)]
	\item the branch transfers control to another basic block within the current trace; or
	\item the branch transfers control to an already translated basic block.
\end{inparaenum} However, not all basic blocks can be connected at translation time, therefore Granary employs edge code and hot-code patching (\Cref{fig:direct_edge_code}) to connect basic blocks at runtime.

Granary atomically patches direct CTIs that target edge code to instead point to their (now resolved) target basic block. Due to limitations of the x86-64 architecture, atomic operations cannot straddle cache lines. Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.

One approach to change an instruction's alignment is to pad the basic block with additional \texttt{nop} instructions. An alternative when more than a few bytes of padding are needed is to introduce a single control-flow instruction (\texttt{jmp}) that jumps around the padded area. However, adding extra instructions introduces latency into the processor's fetch/decode/execute cycle. Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment. For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86-64 processors \cite{AgnerMicroarchitecture}. When possible, alignment requirements are reduced by allowing a patchable instruction to straddle a cache line, but for the part to be patched to be fully contained by a single cache line. Finally, most basic blocks are small (i.e. fit in one cache line), therefore, aligning basic blocks on cache line boundaries naturally avoids the potential for a patchable instruction to straddle two cache lines.

\subsubsection{Trace Allocation Optimization}\label{sec:trace_alloc}
Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}. Typical approaches to tracing require a runtime profiling infrastructure \cite{DynamoRIOOptimization}. Granary attempts to reap the benefits of tracing without the runtime profiling overhead by opportunistically forming traces using ahead-of-time translation (\Cref{sec:aot}), and by arranging for related basic blocks to be allocated sequentially in memory.

Two basic blocks $B_1$ and $B_2$ are ``related" ($B_1 \to B_2$) if there is a control flow instruction from $B_1$ to $B_2$. Granary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions. The base case of this induction are kernel entrypoints. For example, each system call is an entrypoint into the kernel, and is associated with a private memory pool. When a system call is executed for the first time, the basic blocks needed to fulfill the system call's computation are allocated from that system call's memory pool.

To avoid translation conflicts (e.g. on utility functions being simultaneously executed by distinct system calls), and to improve the predictability and reproducability of a given code cache layout, Granary can delay the takeover of kernel entrypoints. That is, Granary can take over one entrypoint at a time, and wait for a specified period of time before taking over the next entrypoint. The effect of delaying the takeover process is that it increases the chances that a given entrypoint (e.g. a blocking system call) will execute to completion, and thus have all of its basic blocks  arranged sequentially in memory.

Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's meta-data (\Cref{sec:metadata}). When Granary needs to translate a new basic block, it inspects the predecessor basic block's meta-data and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.

\subsubsection{System Call Table Optimization}

Granary treats indirect \texttt{jmp}s or \texttt{call}s through the Linux kernel system call table specially. The Linux kernel (as of Linux 3.12) has approximately 300 distinct system call targets, some of which are executed often. To avoid polluting the indirect branch lookup hash table with these targets (\Cref{sec:ibl}), and to avoid the indirect branch lookup overhead, Granary creates a separate, \emph{shadow} system call table.

Granary's translation post-processing mechanism identifies indirect control flow through the system call table and alters those CTIs to direct control flow through the shadow system call table instead. The entries of the shadow system call table are ``lazy" entrypoint translator functions. Upon their first execution, these entrypoint translators find or translate the first basic block of the system call handler, and update the shadow system call table accordingly.

\begin{table*}[ht!]
\scriptsize
\begin{tabularx}{\linewidth}{| l | >{\raggedright}p{0.3\linewidth} | X |}
\hline
\bf Granularity & \bf Description & \bf Example Tool \\
\hline
Kernel & All kernel code is instrumented. & Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters. This data is useful for debugging an performance analysis. Applying the \toolname{cfg} tool to the whole kernel allows us to perform profile-guided optimization of kernel code, including a form of de-virtualization. That is, most indirect control flow instructions (\texttt{call}, \texttt{jmp}) within the kernel target only one destination. Fast checks can be inlined into kernel code that use direct control transfer instructions when possible, and fall-back on indirect control transfer instructions when necessary.\\
\hline
Module & All code from a specific kernel module or driver is instrumented. Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code. All non-module kernel code executes with zero overhead. & Granary's \toolname{Guardrail} tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}. A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs. \\
\hline
Function & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} Linux kernel functions. When invoked by native code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} attaches instrumentation. When invoked by instrumented code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} updates the current nesting depth of the read-side critical section. \\
\hline
\end{tabularx}
\caption{\label{tab:granularities}Different granularities of instrumentation supported by Granary.}
\end{table*}

\section{Only Instrument What is Needed}\label{sec:what}

>>>>

@@

@@

@@

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{Virtual Register Scheduling}

@@

\maketitle
\subsection*{Abstract}

A common challenge faced by static and dynamic binary instrumentation
systems is how to manage transient state used by injected or modified binary
instructions. Virtual register systems solve this problem by providing
instrumentation tool writers with an infinite set of registers for use by
instrumented code. Unfortunately, generic virtual register systems are hard
to implement in practice. In this work, we present a generic virtual
register system for a complex instruction set (x86-64). Our virtual register
system addresses three limitations of prior work: % \cite{Pin,Valgrind}
\begin{inparaenum}[i)]
	\item they require that all code be comprehensively instrumented;
	\item they require a strong notion of threads and the availability of
 thread-local storage, and;
	\item they require a runtime monitor, and therefore cannot be applied
by both static and dynamic binary instrumentation systems.
\end{inparaenum}

%\section{Introduction}\label{sec:intro}
